### 2015-06-02 Final_Ciccone_MS_Thesis.tex
Done

\documentclass{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\newcommand{\Lagr}{\mathcal{L}}
\usepackage{tabulary}
\usepackage{float}
\floatstyle{boxed}
\restylefloat{figure}
\usepackage[section]{placeins}

\begin{document}

\title{A Simple Model of Energy Transition Under a Carbon Budget} 
\author{Adriana Ciccone}
\date{\today}

\maketitle

\begin{abstract}
I develop a model of electricity generation transition from high emitting capital, such as coal or natural gas power generation, to low emitting capital such as solar PV or wind turbines under an exogenous power demand and a carbon dioxide emissions constraint. The model allows early retirement of capital assets, but optimizes by minimizing additional capital built over the length of the simulation. Capital is dependent on investment in high and low emitting capital and an exponential depreciation schedule. I extend the model by allowing capital vintages and including saved operating costs in the objective function. I run simulations over the speed of transition of the productivity of clean energy assets as well as the total amount of $CO_2$ allowed to be emitted. While this model is intended to reflect the existing grid assets and costs, with estimates taken from the EIA, it is not intended to be a detailed prediction of the power sector. Future work can extend this model by endogenizing certain parameters and adding uncertainty in future costs of capital.
\end{abstract}

\section{Introduction}
\subsection{Motivation}

\paragraph{} While there is emerging political consensus in the United States that climate change is occurring and will have significant economic and social impacts, there remain significant barriers to a low-carbon economy. Electric power generation accounts for roughly 30\% of total greenhouse gas emissions in the United States, the largest of any sector \cite{ElectricGHGEmit}. While there now exist renewable power generation sources, such as wind turbines and photovoltaic solar panels, costs remain high compared to traditional coal and natural gas power plants. However, there is some reason to believe that costs will decrease over time as technology improves and economies of scale begin to take effect. With scientific consensus to limit further greenhouse gas emissions to limit warming to an acceptable level, the problem that results is that of energy transition from $CO_2$ intensive to renewable technologies.

\paragraph{} Reducing emissions can be accomplished by replacing durable fossil fuel infrastructure. Existing infrastructure, however, may still have usable life remaining, and discarding it by early retirement is wasteful, considering that there would need to be additional investment in new capital to meet total energy demands. However, in the event of an international treaty or other domestic regulation limiting greenhouse gas emissions, such early retirement may become necessary. In this paper i propose a simple model of a low-cost transition to clean power generation with exactly these constraints; a fixed energy demand and a carbon budget.

\paragraph{} The model proposed on the following pages seeks to answer the question of how best to transition to clean power generation. This model is intentionally simplified to the most basic of descriptive equations to highlight the relationship between key parameters and the outcome. It is not intended to be an exact estimate of the optimal profile of clean energy investments, nor give an exact value of the amount of capital that needs to be retired early. Instead I write down the simplest possible model that will allow us to develop an intuition for the key drivers of capital accrual and early retirement.

\paragraph{} There are two sectors that produce electricity in this model, high-emitting or dirty capital such as coal plants and natural gas generators. The other sector is low-emitting or clean generating capital such as wind turbines and photovoltaic solar panels. Wind and solar are more expensive to build than gas or coal plants for the same unit of energy produced. However, neither wind nor solar emit the  per-kilowatt-hour amount of carbon dioxide that the dirty plants do, and thus will not be subjected to any international or domestic regulation limiting greenhouse gas emissions.

\paragraph{} The optimization requires minimizing the cost of building new capital in both sectors while meeting the domestic energy demand and complying with an emissions cap that is determined exogenously. In addition, existing capital depreciates exponentially regardless of early retirement. This creates an opportunity to substitute clean power generation without having to early retire any coal or natural gas plants. Therefore we could gradually replace high-emitting capital with clean as the high-emitting stock depreciates naturally. However in certain cases, the rate of natural depreciation may not be sufficient to avoid exceeding the emissions cap. In these cases substitution of fully depreciated capital with clean capital would be required, but early retirement of high-emitting capital would also be required. Additionally, the rate at which the cost of clean capital decreases may affect the optimal investment profile, if it becomes feasible that clean energy capital can be cost competitive with existing coal and gas plants. To this effect, I develop this model and run simulations with a range of parameters to show the sensitivity of the optimal transition cost to each of these parameters.

\subsection{Prior Work}

\paragraph{} Prior work in this area has followed two main approaches. Work by Jacobson \cite{Jacobson} seeks to match energy demand and supply using only "wind, water and solar" power, but neglects the optimal transition timing or the cost of early retiring existing capital. Projects such as the National Energy Modeling System attempt to forecast the impact of learning by doing on capital costs and incorporate this in to their projections. However their model is complex, with 20 power-generating technologies. and capital is retired only when it is no longer cost competitive with other generation capacity \cite{NEMS}.

\paragraph{} Jaccard and Rivers attempt to address this problem in their 2006 paper \cite{JR2006}. In it, they create an optimal stopping problem with existing capital stock depreciating linearly and an emissions constraint. They then find the time in the future where the emissions budget has been reached and the entire system must convert to clean technologies. As I explain in the following sections, while their model is simple and develops some intuition for the important of relative capital costs, I extend it by allowing clean and dirty capital to coexist at any given point and find the lowest cost transition profile to meet a similar emissions budget.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Modeling Choices}
\subsection{Model Description}\label{sec:ModelDesc}

\paragraph{} The region or country of interest uses generating capital to produce electricity by two methods: clean (low emitting, represented by $k^l(t)$) and dirty (high emitting, represented by $k^h(t)$). Clean capital produces electricity with low carbon dioxide emissions, but, as a newer technology, it is more expensive per unit of power output than conventional fossil fuel-based generation. Since it is a new technology, the capital cost per kilowatt is expected to decrease over time as manufacturing processes mature and technological innovation produces more efficient designs. High emitting capital is a mature technology that is cheap to build but emits a higher level of carbon dioxide per unit of electricity produced than clean. Since it is a mature technology, there is little reason to believe that production costs will decline substantially in the future.

\paragraph{} Each type of capital depreciates exponentially over time. The decision variables for the optimization are additional investments or disinvestments in each type of capital, represented by four functions ($H^+(t)$, $H^-(t)$, $L^+(t)$, \& $L^-(t)$). These decision variables define the optimal path of investment in dirty (high emitting) and clean (low emitting) capital. All four functions are non-negative over their domain. The functions $H^+(t)$ and $L^+(t)$ represent additional dollars of investment to build new capital of each type. $H^-(t)$ and $L^-(t)$ remove capital from the economy via early retirement. There is no scrap cost to early retirement - the economy simply loses the ability to produce electricity from that particular facility. As I show in the next section, this has consequences for the ability of the economy to meet it's total electricity demand. The retired plant cannot be re-started. Any need for additional generating capital in time t must be met by a positive investment in either $H^+(t)$ or $L^+(t)$.

\begin{equation}\label{eq:simpleHCapitalConstraint}
k^h(t) = H_0 e^{-t/n^h} + \int_0^t e^{x-t/n^h} H^+(x) dx - \int_0^t H^-(x)dx \geq 0 \qquad \forall t
\end{equation}

\begin{equation}\label{eq:simpleLCapitalConstraint}
k^l(t) = L_0 e^{-t/n^l} + \int_0^t e^{x-t/n^l} L^+(x) dx - \int_0^t L^-(x)dx \geq 0 \qquad \forall t
\end{equation}

These equations express total capital in each time period as a function of some exponential depreciation and all prior investments. At each time t, the initial capital endowment, $H_0 = H^+(0)$ and $L_0 = L^+(0)$ depreciate exponentially and any further positive investments depreciate based on their age (x-t). The integral in the above expression shows how each investment $H^+(t)$ or $L^+(t)$ is weighted by an exponential term of the difference in that investment's age and the current time period. Therefore we see that early investments have a more negative exponential term and count less towards existing capital than more recent investments. The rate of depreciation is scaled by $n^h$ and $n^l$, which allows us to vary the lifetime of each type of capital. While we have no limits on the amount of capital in each time period other than the non-negativity constraint on the decision variables, a physical limitation remains that we cannot have negative amounts of capital. We therefore restrict both $k^h(t)$ and $k^l(t)$ to be non-negative in all time periods.

\paragraph{} The objective is to minimize total additional positive investment in both high and low carbon intensity capital for the lowest cost transition to clean energy. We minimize over positive investment only to avoid negative investment - early retirement - benefiting the objective. Instead, early retirement forces additional investment later in the simulation due to an energy generation constraint. That is, if we retire a certain amount of capital early and have a shortfall in generation capacity, we must make a positive investment at some later point. This extra generation would be counted as additional positive investment in the below objective function, penalizing early retirement. The sum is discounted by an interest rate so that we minimize over the net present value of investments.

\begin{equation}\label{eq:simpleObjective}
C(H^+(t), L^+(t)) = \int_0^N (H^+(t) + L^+(t))e^{-rt}) dt \quad if H(t), L(t) \geq 0
\end{equation}

\paragraph{} Note that in this objective function, there is no benefit to early retirement in the form of reduced future operating costs. In the case of dirty technologies, investing in new capital involves pre-purchasing all the required coal or natural gas, as well as labor, operating and maintenance costs. When a plant is retired early, any remaining fuel or variable operating costs don't get redeemed, biasing the model towards investing in clean technologies earlier by penalizing early retirement more than a real-world scenario would. We address this problem in the extended model in section \ref{VintExplan}.

\paragraph{} The economy has some total energy demand that can follow any exogenous path over time. This represents total annual electricity demand, in billions of kilowatt-hours, and therefore daily and seasonal fluctuations in power demand are ignored. Capital must be used to produce electricity to meet this demand, and we must convert from dollars of capital to kilowatt-hours of electricity produced per year to take in to account the relative cost difference of producing one unit of clean energy versus one unit of dirty energy.

\begin{equation}\label{eq:simpleGenConstraint}
G(t) = F^h(t) k^h(t) + F^l(t) k^l(t)
\end{equation}

In this equation, $F^h (t)$ and $F^l (t)$ are the overnight capital cost per billion kilowatt-hours per year, plus the present value of expected future operating costs. That is, they represent the number of kilowatt-hours per year that each dollar of capital can produce. In early years when the clean technology is still relatively new and expensive, a dollar of capital may not produce much energy per year, but a dollar of clean capital invested later in the model may produce significantly more energy. It is important to be clear however, on what this cost is. It is not an annual maintenance cost or any measure of financial obligation for additional production. It is simply the the ability of that unit (dollar) of capital to produce a certain amount of electricity.

\paragraph{} An emissions cap is negotiated by the region's government. Over the course of a fixed number of years, the region must not emit more than a negotiated percent of business as usual emissions. In practice, the government could negotiate to agree to a limit on total number of tons of $CO_2$ emitted instead of negotiating a percentage of business as usual. However, it is convenient to represent a bulk emissions cap as a percent ($\alpha$) reduction for both modeling and to understand how our optimal cost will change as a function of total allowed emissions.

\begin{equation}\label{eq:simpleMaxEmission}
E_{max} = \alpha \int_0^N m^h(t) F^h(t) H^+(0) + m^l(t) F^l(t) L^+(0) dt \qquad \alpha \in [0,1]
\end{equation}

To arrive at an emissions cap, we transform dollars of capital to kilowatt-hours per year as above, then multiply by the carbon intensity of energy, $m^h(t)$ and $m^l(t)$. This value is a measurement of the pounds of $CO_2$ emitted from power generation per kilowatt-hour. We can then write actual emissions over the length of simulation as the following.

\begin{equation}\label{eq:simpleSimEmission}
E(H, L) = \int_0^N  m^h(t) F^h(t) H_0 + m^l(t) F^l(t) L_0 dt
\end{equation}

Total emissions over the treaty's length must not exceed the cap represented by a fraction of business as usual, leading to the final constraint:

\begin{equation}\label{eq:simpleEmitConstraint}
E(H,L) \leq E_{max}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

\subsection{Implementation}
The code for the following results was written in Python using the Pyomo modeling and optimization library. All data and code can be found at \\ https://github.com/cicconea/EnergyTransition. Data for parameters comes from the Energy Information Administration (EIA) unless otherwise stated.

\subsection{Functional Form Choices}

\paragraph{} Several key parameters are allowed to vary over time. In this section I explain the modeling choices and sources of data for each of these parameters.

\paragraph{} Energy demand is exogenous and comes directly from the EIA. From the EIA demand forecast, I subtract generating capacity supplied by hydroelectric dams and nuclear plants to evaluate only the subset of demand that will be fulfilled by the clean (wind, solar photovoltaic) and dirty (coal, natural gas) technologies. I consider this an acceptable simplification that can still show the dynamics of the clean energy transition, since hydroelectric dams, barring some small-scale technological innovation, have reached their maximum penetration in the United States, and the future of nuclear is unclear. Future work may want to include hydro and nuclear as well as separate wind, solar, gas and coal in to separate sectors. Energy demand growth is roughly linear over time, going from 2836 billion kWh in 2011 to 3734 billion kWh in 2040. For any simulation length greater than 30 years, I assume that growth will continue to grow at the same per year rate.

\paragraph{} While there may be future improvements in the cost of coal and natural gas plants, we might expect the majority of capital cost per kilowatt-hour decreases to come from the continuing switch from coal to gas-fired generation over the next 50 years. According to the EIA, coal powered generating capacity will go from 1717 billion kWh/year in 2011 (or 43\% of net generating capacity) to 1661 billion kWh/year (34\% of net generating capacity) in 2040, while natural gas will go from 925 billion kWh/year (23\%) in 2011 to 1605 billion kWh/year (33\%) in 2040 \cite{EIATable5}. Also from the EIA are estimates of current overnight capital costs to build a kW of generating capacity \cite{EIACapCost}. I then compute a weighted average overnight capital cost from the projected generating capacity and the current overnight capital costs. Note that $F^h(t)$ is the inverse of this weighted average. To arrive at a generating capacity per year, I convert the kW generating capacity to kWh/year assuming a capacity factor of 50\%. Similarly, I assume a weighted average of carbon intensities of energy for coal and gas. As the US economy transitions to having more gas-fired power plants, the carbon intensity of energy will go down over time. A summary of these parameters can be found in Table \ref{tab:ParamValues}.

\begin{figure}[h]
\includegraphics[width=0.9\textwidth]{Fl_Simulation_Range.png}
\caption{Scenarios of clean capital cost transition speed \label{fig:FlScaleScenarios}}
\end{figure}

\paragraph{} Of course, the main source of uncertainty in this model is the speed of cost reductions for clean energy capital. I model this forecast deterministically as a logistic function, with some initial low value and a maximum number of kWh/year per dollar that represents the cost floor of a mature technology. Varying the speed of transition between the minimum and maximum kWh/year/dollar, as well as maximum output/\$ value allows us to understand the sensitivity of the objective function with respect to the speed by which clean capital cost decreases. Figure \ref{fig:FlScaleScenarios} gives some cost curve scenarios in comparison with the linear high-emitting capital cost curve. The minimum cost in each of these figures is half of the current weighted average overnight capital cost from the EIA and the maximum is the cost per kWh/year of natural gas. Both the minimum and the maximum are then adjusted to account for a lower renewable capacity factor (30\%) that accounts for the intermittent nature of renewable power generation. Emissions per kWh of clean energy are assumed to be 0; that is, any $C0_2$ emissions from the construction of the wind turbines or photovoltaic cells are assumed to be negligible.

\FloatBarrier

\begin{table}[h]\label{tab:ParamValues}
\begin{tabulary}{1.0\textwidth}{LLLL}
Variable & Description & Value\\
\hline
$\alpha$ &  \% Business As Usual Emissions  &  Varied \\
$Fl_{min}$ & Minimum kWh/year per dollar - Clean Capital & 0.343349882\\
$F_{max}$ & Maximum kWh/year per dollar - Clean Capital & 2.8658669 \\
$Scale$ & Transition Speed  (Scale of Logistic Function) & Varied \\
$Fh_{min}$ & Minimum kWh/year per dollar - Dirty Capital & 2.628 \\
$Fh_{max}$ & Maximum kWh/year per dollar - Dirty Capital & 4.7764449 \\
$G_0$ & Base energy demand (kWh/year) & $2798.5 * 10^9$ kWh\\
$G_m$ & Annual increase in energy demand (kWh/year) & $32.238 * 10^9$ kWh \\
$c^l$ & Capacity Factor Clean Capital & 30\% \\
$c^h$ & Capacity Factor Dirty Capital & 50\% \\
\hline
\end{tabulary}
\caption{Key parameters of simulation. Note that energy demand is excluding the generating capacity of hydroelectric dams and nuclear plants}
\end{table}

As we can see from the parameters in this table, clean capital has a higher cost per kWh/year than dirty capital, yet it has a significantly lower carbon footprint. By forcing capital in each period to meet some exogenous energy demand, we force the model to reconcile these two competing parameters, leading to changes in optimal investment profiles that fulfill the modeling requirements.

\FloatBarrier

\subsection{Model Results}
\FloatBarrier

\paragraph{} I run the model multiple sets of parameter values. Specifically, three parameters of interest are the speed of clean capital cost reductions (Scale in table \ref{tab:ParamValues}), the minimum cost reached by low emitting capital, and the fraction of business as usual emissions that define the emissions cap. Based on the equations in section \ref{sec:ModelDesc}, we would expect the minimum cost of the transition to increase with the increasing restriction of the emissions cap. As we require a smaller and smaller percentage of $CO_2$ to be emitted over the length of the simulation, we must build more and more low-emitting (clean) capital. Since it is more costly to build per kilowatt-hour than coal or gas plants, our objective function is penalized. Figure \ref{fig:costVsAlpha} shows the relationship between emissions reduction and total transition cost.

\begin{figure}[h]
\includegraphics[width=\textwidth]{cost_as_func_of_Alpha.png}
\caption{Cost vs Alpha\label{fig:costVsAlpha}}
\end{figure}

\FloatBarrier

\paragraph{} For high values of $\alpha$, the emissions cap is small compared to business as usual - that is, we commit to 1-$\alpha$ reduction in BAU emissions. Consequently, we do not have to build large amounts of clean capital to accomplish such a reduction and because it is less cost-efficient to build, we chose not to invest heavily in clean technologies. Figure \ref{fig:SimpleResultsA=50} gives an example of such a transition where the emissions cap is set to emit no more than 50\% of business as usual emissions.

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_5_alpha_50_FLFrac_50.png}
\caption{A slow transition speed with a lenient emissions reduction fraction and moderate of renewable capital\label{fig:SimpleResultsA=50}}
\end{figure}

\FloatBarrier

\paragraph{} In comparison, when the emissions cap is set much lower, say 25\% of business as usual emissions, investment in clean technologies must happen much sooner and with a higher magnitude of investment. Such a transition is shown in figure \ref{fig:SimpleResultsA=25}. An increased transition speed means that the cost of clean has come down faster than in Figure \ref{fig:SimpleResultsA=50}, resulting in a lower total capital investment. In the figure below, the investment spike in clean energy due to early retirement around year 27 is roughly half of the corresponding investment made in the prior figure as a result. This spike in investment is driven by the early retirement of polluting capital to avoid exceeding the emissions cap through the remainder of the simulation. Thus we can see a stricter emissions limit in the figure below, where emissions were limited to 25\% of business as usual, results in an earlier timeline for early retirement than in Figure \ref{fig:SimpleResultsA=50}.

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_5_alpha_25_FLFrac_50.png}
\caption{Slow transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:SimpleResultsA=25}}
\end{figure}

\FloatBarrier

%\paragraph{} The characteristic spike in clean capital investment appears in all but the most extreme cases of parameter selection. Shown below are results from running the simulation over various speeds of clean capital cost reduction. As we would expect, the faster the cost reduction speed, the lower the cost of transition because we can take advantage of more cost-effective clean capital earlier in the simulation, reducing the need to early retire polluting capital once the economy nears it's emissions cap.

\paragraph{} Similarly, the fast cost transition speed scenario shown below again shows a reduction in the total clean capital required to replace the same amount of polluting technology being early retired. Since the clean capital is more productive in the fast transition scenario, fewer dollars are needed to meet the generating demand than in figure \ref{fig:SimpleResultsA=50}.

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_15_alpha_50_FLFrac_50.png}
\caption{Fast transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:SimpleResultsFl=1}}
\end{figure}

\FloatBarrier

Finally, varying the final cost per kilowatt-hour of the clean capital has an effect on the transition profile from polluting to clean technologies. In figure \ref{fig:SimpleResultsFl=1}, we can see that since the clean capital is more productive, since it reaches a higher kWh/\$ in each year than a similar cost curve with the same transition speed, but a lower final cost. Therefore while the year of early retirement remains the same, the total amount of clean capital required to be invested is less because each unit is more productive.

\begin{figure}[h]
\includegraphics[width=\textwidth]{simple_cap_and_invest_results_Fl_15_alpha_50_FLFrac_100.png}
\caption{Fast transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:SimpleResultsFl=1}}
\end{figure}

\FloatBarrier

\subsection{Vintaged Model}\label{VintExplan}

\paragraph{} In the simple model described above, the mechanics of the problem are explained simply and with supporting intuition. However this simple version ignores some key facts about the energy system that are relevant to the solution. The first simplification is that all capital of a particular type in a certain time has the same kilowatt-hour/year per dollar value ($F^h(t), F^l(t)$). That is, the investment of any capital in an early period benefits from improved cost effectiveness in later periods despite being physically the same plant. Treated as a lump sum of capital, this simplification is understandable, yet it biases results towards premature investment in a technology by assuming that any early investment will benefit from a decreasing future cost curve.

\paragraph{} To address this issue, we introduce capital vintages. In each year of the simulation we allow investment in high and low emitting capital. In subsequent years this capital depreciates exponentially but can also be early retired, as long as the total capital of that vintage remains a non-negative number. Then we define our kilowatt-hour per dollar value profile based on the vintage year, not the simulation year so that we can track the accurate generating capacity of the whole sector. This creates an incentive in the model to early retire old vintages of capital first, as they provide the least generation capability per dollar, mimicking a rational investment decision that we would expect in reality.

\paragraph{} The second simplification I seek to address is the lack of benefit to early retirement from saved operating costs. In the simple model described above, all capital and operating costs are paid up front in the investment cost. If a certain amount of capital is retired before the end of it's useful life, the remaining years of operating costs cannot be redeemed. This biases the results above by penalizing early retirement and spurring earlier investment in clean capital.

\subsection{Model Math}

\paragraph{} Below are the equations describing a model with vintaged capital stock. In each year, i, we allow some non-negative investment in either high or low emitting capital. In all subsequent periods, this initial positive investment depreciates exponentially according to it's age (t-i). We also allow early retirement in all future periods as long as capital of either type remains non-negative in every time period.

\begin{equation}
k_i^h(t) = H_i^+ e^{i-t/n} - \int_i^t H^-_i(x)dx
\end{equation}

\begin{equation}
k_i^l(t) = L_i^+ e^{i-t/n} - \int_i^t L^-_i(x)dx
\end{equation}

\begin{equation}
H^+_i, H^-_i(t), L^+_i(t), L^-_i(t) \geq 0 \forall i \in [0,N], t \in [i,N]
\end{equation}

\paragraph{} The objective function for this model is slightly more complex as it takes in to account the benefit of early retirement in the form of reduced future operating costs. If we let $\beta^h, \beta^l$ be annual operating costs for high and low emitting capital stocks, then without early retirements the present value operating costs for vintage $i$ of high emitting capital would be:

\begin{equation}
OC^+_i=\beta \int_i^N H_i^+e^{i-t/n}e^{-rt}dt
\end{equation}

\paragraph{} The actual present value of operating costs for a particular vintage is the capital of vintage i at time t, multiplied by the operating cost $\beta^h$ and discounted by an interest rate r, with an analogous expression for the low emitting capital stock.

\begin{equation}
OC^{actual}_i=\beta \int_{t=i}^N \left[H_i^+e^{i-t/n}dt -\int_i^tH_i^-(x)dx\right] e^{-rt}dt.
\end{equation}

\paragraph{} The difference between these two terms is the savings in operating costs from early retirement for vintage $i$. Substituting this expression for both high and low emitting capital and integrating over all vintages  i $\in [0,N]$:

\begin{equation}
Savings = \beta^h \int_{i=0}^N \int_{t=i}^N \int_{x=i}^t H^-_i(x) dx e^{-rt} dt di + \beta^l \int_{i=0}^N \int_{t=i}^N\int_{x=i}^t L_i^-(x) dx e^{-rt} dt di
\end{equation}

\paragraph{} The objective function is then the present value of installation costs minus any savings from early retirement:

\begin{equation}\label{eq:VintObj}
C = \int_{i=0}^N  \left[ (H^+_i + L^+_i)e^{-ri}  - \beta^h  \int_{t=i}^N \left[\int_{x=i}^t H^-_i(x) dx e^{-rt}\right] dt - \beta^l  \int_{t=i}^N \left[\int_{x=i}^t L^-_i(x) dx e^{-rt}\right] dt\right] di
\end{equation}

\paragraph{} We then minimize this net cost with subject to the same generation constraints and emissions constraint as before. Note that both kilowatt-hours/year per dollar ($F^h(i), F^l(i)$ and carbon intensity ($m^h(i), m^l(i)$) now change by vintage, not by year. That is, capital built in year one will have the same productivity and carbon intensity throughout it's life, regardless of how old it gets, though depreciation will lead to a reduction in capital available to produce power.

\begin{equation}
G(t) = \int_0^t F^h(i)k_i^h(t) + F^l(i)k_i^l(t) di
\end{equation}

\begin{equation}
\int_{t=0}^{t=N}\int_{i=0}^{i=t} m^h(i)F^h(i)k_i^h(t) + m^l(i)F^l(i)k_i^l(t) di dt \leq \bar{E}
\end{equation}

\subsection{Results}

\paragraph{} Using the same assumptions about the functional forms of the productivities of capital and their emissions potential, I now implement and run a version of the same model with vintages of capital. Note that now each $F^h$ and $F^l$ now correspond to the productivity of each vintage, not of time. That is, in each period of the simulation the model can invest in new capital at that period's productivity and if, in future periods some of that capital is retired early, the generating capacity of the system is reduced by that vintage's productivity and not the productivity of the period in which it is retired.

\paragraph{} The results of the vintaged simulation are shown below for the same set of parameters as the simple model results shown above. Though the parameters for each simulation are the same, some important differences emerge in the results of the vintaged model. In figure \ref{fig:VintResultsFl=5_Alpha_50_Frac50}, early retirement of fossil fuel based capital still occurs, however it retires later, in year 30 instead of in year 27. Smaller amounts of high-emitting capital are retired, and are replaced by a smaller amount of clean capital. Remarkably, however, the total profile of the capital transition are identical, though the vintaged model requires more total capital in each year.

\begin{figure}[h]
\includegraphics[width=\textwidth]{vint_cap_and_invest_results_Fl_5_alpha_05_FLFrac_05.png}
\caption{Slow transition speed with 50\% emissions reduction fraction and moderate cost of renewable capital\label{fig:VintResultsFl=5_Alpha50_Frac50}}
\end{figure}

\FloatBarrier

\begin{figure}[H]
\includegraphics[width=\textwidth]{vint_cap_and_invest_results_Fl_15_alpha_05_FLFrac_05.png}
\caption{Fast transition speed with 50\% emissions reduction and moderate cost of renewable capital\label{fig:VintResultsFl=15_Alpha50_Frac50}}
\end{figure}

\paragraph{} Speeding the transition speed of clean technologies while keeping all other parameters equal has the same effect as in the simple model; early retirement is delayed as shown in figure \ref{fig:VintResultsFl=15_Alpha50_Frac50}. And finally, shifting the emissions cap to be more strict results in any early retirement happening much earlier in the simulation, along with an earlier investment in clean capital, as shown in figure \ref{fig:VintResultsFl=15_Alpha25_Frac50}

\begin{figure}[H]
\includegraphics[width=\textwidth]{vint_cap_and_invest_results_Fl_5_alpha_025_FLFrac_05.png}
\caption{Slow transition speed with 25\% emissions reduction and moderate cost of renewable capital. \label{fig:VintResultsFl=15_Alpha25_Frac50}}
\end{figure}

\paragraph{} Note that in the scenario shown in figure \ref{fig:VintResultsFl=15_Alpha25_Frac50}, no additional high-emitting capital is built, and in fact some must be early retired in order to meet the strict emissions requirement. This is a scenario where the emissions requirement is so stringent that we must only build clean capital over the entire simulation. This implies that for certain parameter values, the United States may already be over-built in polluting capital.

\paragraph{} In all of these vintaged model, early retirement brings a benefit of savings on operating costs. This would drive the model to early retire earlier and larger amounts than in the simple model, though we do not observe this. This indicates that the effect of the emissions constraint dominates the benefit of operational savings.

\FloatBarrier

\section{Model Discussion}

\paragraph{} While the input parameters to the model were generated to only roughly replicate the US power market, we can see that nearly regardless of the speed of transition of clean capital we will have to transition a significant portion of the existing generation to renewable technologies. A further cost, which I do not consider here, is the durable infrastructure for fuel delivery to the plants. Natural gas pipelines, rail lines for coal, and delivery trucks for coal ought to all be considered as initial capital that will depreciate over time, and their operating costs such as maintenance, labor, and fuel requirements could be included in an objective function such as \ref{eq:VintObj}. These costs are secondary effects of the model suggesting to retire high-emitting capital early, so any capital retirement will have knock on effects to this infrastructure by means of lost volume. Of course, this model makes no mention of the power distribution system. While the infrastructure leading from natural gas wells and coal mines up to the power plants may change dramatically in certain modeling scenarios, I assume that the existing distribution infrastructure is not changing as a result of the production system changes. Demand increases would drive distribution system growth, and depreciation of these assets would be managed as part of any utility's planned process. While this model does not attempt to specify where in the country solar or wind assets would be placed, the cost of changing the transmission system may only be relevant to the costs in these models if a significant amount of distributed generation capacity from solar panels or remote wind farms requiring additional transmission lines to be built.

\paragraph{} This model also simplifies the power generation sector in to two generation sources; clean and dirty. While this solution still provides the correct intuition for a transition to renewable energy, it neglects the unique cost characteristics of each technology in consideration. A model that separates each sector may lead to a better understanding of the timing and limitations of each type of clean energy technology. However with each additional parameter that must be estimated, another source of error is included in the model which may mask the true opportunities or challenges of the clean energy transition.

\paragraph{} Of course, parameter estimation is not without concern in the existing model. I make some crude estimates for the minimum and maximum kilowatt-hours/year per dollar of wind and solar panels, however there may be reasons for which the cost of clean capital becomes cost competitive with gas or coal due to improvements in technology or manufacturing capabilities. Conversely, there may be limitations to how cheap clean capital can become, particularly if such an international treaty were reached and all countries were making this transition, driving the price of solar panels or wind turbines up.

\begin{figure}[h]
\includegraphics[width=\textwidth]{cost_as_func_of_Fl_emissions_05.png}
\caption{Cost vs Fl\label{fig:costVsFl}}
\end{figure}

\paragraph{} Specific peculiarities of the power generation sector are not modeled here. For example, we make no assumptions about the frequency of running a plant that arises from the bid-in nature of marginal cost power generation. The simplifying assumption is that a capacity factor included in the kilowatt-hours/year per dollar of capital addresses this by combining the profile of frequency of generation coupled with the intermittency of renewable technologies.

\paragraph{} However the most important sensitivity of the model is to the speed of clean capital cost reductions. Shown below in figure \ref{fig:costVsFl} are the optimal cost solutions to simulations run with different cost reduction speeds. A higher value on the x-axis means that the speed of transition is faster. Examples of the range of transmission profiles can be found above in figure \ref{fig:FlScaleScenarios}. For slow transitions, the model is extremely sensitive to improvements in cost reductions. As the speed of transition increases, the model becomes less sensitive to improvements, presumably because we reach the minimum cost per kilowatt-hour quickly. A natural extension to this model would be endogenizing the cost of clean capital using a learning by doing approach. Alternatively, future work could include modeling this problem with uncertainty in the future distribution of cost reductions. Currently the model has perfect foresight, which may lead to over-investment in renewable technology early in the simulation as planners know that the cost effectiveness will increase in the future. With uncertainty in the cost of clean capital, we would better reflect the current challenges facing the clean energy transition

\section{Conclusion}

\paragraph{} This project was motivated by the desire to understand the effects of key parameters on a transition to clean energy technologies with the constraint of a maximum emissions level. I set up a model that extended a version of Jacquard and Rivers's 2006 paper and performed a sensitivity analysis over key values of parameters. While the model does not attempt to mimic the details of the power sector in the United States, however it demonstrates the consistency of the optimal profile of transition, including early retirement of existing fossil fuel-based stocks in all but the most forgiving of scenarios. Furthermore, the model is highly dependent on the lifetime of existing capital stock. Instead of modeling lifetimes of capital as dependent on the price of electricity it supplies, I model a fixed depreciation schedule for each vintage. If this lifetime changes, or if the model is changed to assume a fixed operation lifetime followed by shutdown, the profile of early retirements may change dramatically. Specifically for longer-lived assets, the amount of capital that will need to be retired early may increase significantly.

\paragraph{} I extend the simple model developed above by including vintages of capital and including the savings from operating costs. While some small changes to the results occur by including this complication, ultimately including vintages does not change the model profile or results significantly. There are other extensions that should be considered, however. First, investigating the sensitivity of the model results to varied depreciation schedule of the assets could be modeled more realistically. Endogenizing capital asset productivity using a learning by doing approach may also change results by incorporating productivity increases when clean technologies are invested in earlier than the existing model recommends. Finally, adding uncertainty over the future costs of clean technologies will better mimic the real world situation where we do not have perfect foresight over the costs of solar and wind power over the next century.

\pagebreak
\appendix
\section{First Order Conditions for Simple Model}

In order to attempt an analytical solution, I make several key assumptions about the functional forms of key parameters. While the equations for capital stocks remain the same, I assume a flat, fixed energy demand, $G$, through the length of the simulation. The energy per unit of capital for high-emitting technologies is similarly fixed at $F^h$ for the entire period, while it decays exponentially for low-emitting technology with a half life of $\tau$. Assuming these functional forms is arguably a good approximation of existing beliefs about the future costs of low-emitting technologies with learning by doing or economies of scale; that is, a rapid improvement in the output per dollar of capital that slowly levels off as t grows large. I further assume that high-emitting carbon intensity is fixed at $m^h$ and that there are no emissions associated with clean capital. With these simplification we now write the following optimization problem. \\

Objective: Chose $H^+(t)$, $H^-(t)$, $L^+(t)$, and $L^-(t)$ to minimize:

\begin{equation}\label{eq:analyticalObj}
C = \int_0^N (H^+(t) + L^+(t))e^{-rt} dt
\end{equation}

Subject to:

\begin{multline}\label{eq:analyticalGen}
G = \bar{F}^h\left (H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \right) \\ + \bar{F}^l(1-e^{-t/\tau}) \left ( L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx\right )
\end{multline}

\begin{equation}\label{eq:analyticalEmit}
m^hF^h\left (H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \right) \leq \alpha N m^h F^h H^+(0)
\end{equation}

\begin{equation}\label{eq:analyticalHCapConstraint}
H_0 e^{-t/n} + \int_0^t e^{x-t/n} H^+(x) dx - \int_0^t H^-(x)dx \geq 0 
\end{equation}

\begin{equation}\label{eq:analyticalLCapConstraint}
L_0 e^{-t/n} + \int_0^t e^{x-t/n} L^+(x) dx - \int_0^t L^-(x)dx \geq 0 
\end{equation}

\begin{equation}\label{eq:analyticalDecisionVarConstraint}
H^+(t), H^-(t), L^+(t), L^-(t) \geq 0
\end{equation}

The first order conditions for a solution to the model are listed below in equations \ref{eq:dLdH+} through \ref{eq:dLdpi}.

\begin{equation}\label{eq:dLdH+}
\frac{d\Lagr}{dH^+(t)} = 1/r(e^{-rN} -1) + n(1-e^{-t/n})(\rho - \lambda F^h -\mu m F^h) + \sigma = 0
\end{equation}

\begin{equation}\label{eq:dLdH-}
\frac{d\Lagr}{dH^-(t)} = (\lambda F^h +\mu m F^h - \rho)t + \phi = 0
\end{equation}

\begin{equation}\label{eq:dLdL+}
\frac{d\Lagr}{dL^+(t)} = 1/r(e^{-rN} -1) + n(1-e^{-t/n})(\pi - \lambda F^l(1-e^{-t/\tau})) + \gamma = 0
\end{equation}

\begin{equation}\label{eq:dLdL-}
\frac{d\Lagr}{dL^-(t)} = (\lambda F^l(1-e^{-t/\tau}) - \pi)t + \delta = 0
\end{equation}

\begin{multline}\label{eq:dLdlambda}
\frac{d\Lagr}{d\lambda} = G - F^h\left[nH^+(t) + e^{-t/n}(n^2 + H_0(1-n)) - n^2 + \int_0^tH^-(x)dx\right] \\ - F^l (1-e^{-t/\tau})\left[nL^+(t) + e^{-t/n}(n^2 + L_0(1-n)) - n^2 + \int_0^tL^-(x)dx\right] = 0
\end{multline}

\begin{multline}\label{eq:dLdmu}
\frac{d\Lagr}{d\mu} = \bar{E} - mF^h \left [(1-e^{-N/n})(1-n)H_0 n + n^2(N + n(1-e^{-N/n})) \right.\\ + \left. n\int_0^N H^+(t) dt - \int_0^N\int_0^t H^-(x)dx dt   \right] \geq 0
\end{multline}

\begin{equation}\label{eq:dLdrho}
\frac{d\Lagr}{d\rho} = H_0 e^{-t/n}(1-n) + nH^+(t) - n^2(1-e^{-t/n}) - \int_0^t H^-(x)dx \geq 0 
\end{equation}

\begin{equation}\label{eq:dLdpi}
\frac{d\Lagr}{d\pi} = L_0 e^{-t/n}(1-n) + nL^+(t) - n^2(1-e^{-t/n}) - \int_0^t L^-(x)dx \geq 0 
\end{equation}

\pagebreak
\addcontentsline{toc}{chapter}{References}

\begin{thebibliography}{99}

\bibitem{Jacobson}
MZ Jacobson, MA Delucchi. Providing all global energy with wind, water, and solar power, Part I: Technologies, energy resources, quantities and areas of infrastructure, and materials. Energy Policy 39 (2011) 1154?1169

\bibitem{NEMS}
Energy Information Administration. The National Energy Modeling System: An Overview. October 2009

\bibitem{JR2006}
Jaccard and Rivers, Heterogeneous capital stocks and the optimal timing for CO2 abatement. Resource and Energy Economics 29 (2007) 1?16.

\bibitem{EIATable5}
Energy Information Administration. Annual Energy Outlook 2014

\bibitem{EIACapCost}
Energy Information Administration. Updated Capital Cost Estimates for Utility Scale Electricity Generating Plants. April 2013

\bibitem{ElectricGHGEmit}
Environmental Protection Agency. Sources of Greenhouse Gas Emissions.  http://www.epa.gov/climatechange/ghgemissions/sources/electricity.html . Accessed 4/25/2015

http://www.eia.gov/forecasts/capitalcost/

\end{thebibliography}

\end{document}